---
title: "Project"
author: "Marco Sorbona"
subtitle: "Investigation into Chicago taxis"
echo: false
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r}
#| warning: false
#| message: true
library(tidyverse)
library(tidymodels)
library(scales)
library(ggthemes)
```

## Data Overview

The `taxi` dataset contains information on a subset of taxi trips in Chicago during 2022.

Below is a snapshot of the dataset:

```{r}
#| label: glimpse
glimpse(taxi)
```

## Question 1: Is there a relationship between trip distance and tipping?

**Research question:** Do passengers tip differently based on how far they travel?

Hypotheses:

-   **Hp1**: Longer trips are more likely to receive a tip

-   **Hp2**: Shorter trips are cheaper, so passengers might tip more frequently

This question involves exploring the relationship between a **continuous variable** (distance in miles) and a **categorical variable** (tip: yes/no).

------------------------------------------------------------------------

### Initial Exploration: Tip Frequency

First, let's look at the overall proportion of trips that received tips:

#### Proportion with Bar plot

```{r}
#| label: q1-proportion-bar-plot
#| fig-cap: "Proportion of trips with and without tips"
taxi |>
  count(tip) |>
  mutate(perc = n / sum(n)) |>
  ggplot(aes(x = tip, y = perc, fill = tip)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Tip proportion in Chicago taxi trips",
    subtitle = "The vast majority of trips receive tips",
    x = "Tip received",
    y = "Proportion of trips",
    fill = "Tip"
  ) +
  theme_minimal()
```

**Observation:** Most trips in this dataset received a tip.

------------------------------------------------------------------------

#### Summary Statistics

```{r}
#| label: summary-stat
taxi |>
  group_by(tip) |>
  summarise(
    mean_dist = mean(distance, na.rm = TRUE),
    median_dist = median(distance, na.rm = TRUE),
    sd_dist = sd(distance, na.rm = TRUE),
    n = n()
  )
```

------------------------------------------------------------------------

#### Distribution of Trip Distances by Tip Status

#### Histogram

```{r}
#| label: q1-histogram
#| fig-cap: "Distribution of trips with and without tips"
#| warning: false
ggplot(taxi, aes(x = distance, fill = tip)) +
  geom_histogram(binwidth = 0.15, alpha = 0.7, position = "identity") +
  scale_x_continuous(limits = c(-0.5, 25), breaks = seq(0,50,5)) +
  scale_y_continuous(breaks = seq(0, 800, 200)) +
  labs(
    title = "Distribution of trip distance by tip status",
    subtitle = "Multimodal distribution with peaks at ~2, ~12, and ~18 miles",
    x = "Distance (miles)",
    y = "Number of trips",
    fill = "Tip"
  ) +
  theme_minimal()
```

The histogram reveals a **multimodal distribution** with **four distinct peaks**:

-   **Peak 1: \~0 miles**. Very short/zero-distance trips

-   **Peak 2: \~2.5 miles**. Most common trip length

-   **Peak 3: \~12 miles**. Medium-length trips

-   **Peak 4: \~18 miles**. Long trips

#### What Could the 0-Mile Peak Represent?

The peak at 0 miles suggests data quality issues (perhaps trips with missing distance), which should be investigated or filtered.

**Possible explanations:**

1.  **Data errors**: Trips recorded with zero distance (technical issues?)

2.  **Cancelled trips**: Passenger cancelled but driver still recorded?

3.  **Very short hops**: Around the block, between nearby buildings

4.  **Airport waiting**: Taxis waiting in queue with zero movement

------------------------------------------------------------------------

#### Boxplot

```{r}
#| label: q1-boxplot
#| fig-cap: "Trip distance distribution by tipping status"
ggplot(taxi, aes(x = tip, y = distance, fill = tip)) +
  geom_boxplot() +
  labs(
    title = "Trip distance distribution by tip status",
    subtitle = "Boxplot shows many outliers at longer distances",
    x = "Tip received",
    y = "Distance (miles)",
    fill = "Tip"
  ) +
  theme_minimal()
```

The boxplot confirms the presence of numerous outliers at longer distances, making it less ideal for visualizing the overall pattern.

------------------------------------------------------------------------

#### Density Plot (Best Visualisation)

```{r}
#| label: q1-density-plot
#| fig-cap: "Density of trip distances by tipping status"
#| warning: false
ggplot(taxi, aes(x = distance, color = tip, fill = tip)) +
  geom_density(alpha = 0.4, linewidth = 0.5) +
  scale_x_continuous(limits = c(0, 25)) +
  labs(
    title = "Trip distance density by tip status",
    subtitle = "Longer trips show higher probability of tipping",
    x = "Distance (miles)",
    y = "Density",
    color = "Tip",
    fill = "Tip"
  ) +
  theme_minimal()
```

### Findings

The density plot reveals several patterns:

1.  **Short trips (\<5 miles)**: Approximately equal density for tipped and non-tipped trips. Passengers are equally likely to tip or not.

2.  **Medium trips (10-15 miles)**: A small peak shows non-tipped trips slightly more common.

3.  **Long trips (\>15 miles)**: Tipped trips show higher density, suggesting passengers are more likely to tip on longer journeys.

4.  **Peak at 0 miles**: Likely data quality issue (trips with zero recorded distance) that should be investigated.

### Conclusion

**There is a clear relationship between trip distance and tipping behavior.**

The density plot reveals that:

1.  **Short trips** (around 2 miles) have roughly equal probability of receiving a tip or not

2.  **Medium trips** (around 12 miles) show a higher probability of **not** receiving a tip

3.  **Long trips** (around 18 miles) show a substantially higher probability of receiving a tip

This pattern supports **Hypothesis 1**: longer trips are more likely to receive a tip. The relationship is not simply linear, there appears to be a threshold around 15 miles where tipping probability increases noticeably.

The density plot is the most effective visualization for this question as it clearly shows probability differences across the distance spectrum

## Question 2: Do taxi passengers tend to tip more for the company Chicago Independents than the other companies?

**Research question**: Do passengers tip at higher rate when riding with Chicago Independents compared to other taxi companies?

**Hypotheses**:

-   Hp0: There is no difference in tip rates between Chicago Independents and other companies

-   Hp1: Chicago Independents receive tips at a higher rate than other companies

This question involves exploring the relationship between a categorical variable (company) and the proportion of trips that received tips.

------------------------------------------------------------------------

### Initial exploration: Calculate tip rate

First, we need to calculate the tip rate for each company: the proportion of trips where a tip was given.

```{r}
#| label: q2-company-tip-rate
#| fig-cap: "Tip rate by company"

company_tip_rates <- taxi |>
  group_by(company) |>
  summarise(
    total_trips = n(),
    tipped_trips = sum(tip == "yes", na.rm = TRUE),
    tip_rate = tipped_trips / total_trips
  ) |>
arrange(desc(tip_rate))

company_tip_rates
```

------------------------------------------------------------------------

#### Highlight Chicago Independents

To make comparison easier, we create a flag for Chicago Independents

```{r}
company_tip_rates <- company_tip_rates |>
  mutate(
    highlight = if_else(company == "Chicago Independents", "CI", "Other")
  )

company_tip_rates
```

------------------------------------------------------------------------

#### Visualisation

A bar plot allows us to compare tip rates across companies, with Chicago Independents highlighted:

```{r}
ggplot(company_tip_rates, aes(x = reorder(company, tip_rate), y = tip_rate, fill = highlight)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("CI" = "red", "Other" = "gray")) +
  labs(
    title = "Do passengers tip more for Chicago Independents?",
    subtitle = "Tip rate = % of trips with a tip",
    x = "Company",
    y = "Tip rate",
    caption = "Chicago Independents highlighted in red"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### Findings

The analysis shows that **Chicago Independents have the highest tip rate** among all companies, at approximately **95%** of trips receiving a tip. This is higher than the tip rate across other companies.

------------------------------------------------------------------------

### Discussion

The higher tip rate for Chicago Independents could be explained by several factors:

-   Better service quality: Drivers may provide a superior experience

-   Operational differences: Perhaps they serve different neighborhoods or longer trips

-   Passenger demographics: Their customer base might have different tipping norms

##### Limitations and follow-up questions

The current analysis cannot explain why Chicago Independents have higher tip rates. Access to more granular data would help investigate:

-   Do they operate primarily in wealthier neighborhoods?

-   Do they perform longer trips on average? (Previous analysis showed longer trips are tipped more frequently)

-   Is there variation in service quality (vehicle condition, driver professionalism)?

-   Do they use different payment methods that encourage tipping?

These questions point to the need for multivariate analysis to disentangle the factors contributing to higher tip rates.

------------------------------------------------------------------------

## Follow-up Question: Do Chicago Independents perform longer trips on average?

#### Rationale

Question 1 established that longer trips are more likely to receive tips. Question 2 found that Chicago Independents have the highest tip rate among all companies (approximately 95%). This follow-up investigates whether the higher tip rate for Chicago Independents might be explained by systematically longer trips, rather than company-specific factors like service quality.

#### Hypotheses

-   H0: There is no meaningful difference in trip distances between Chicago Independents and other companies

-   H1: Chicago Independents perform significantly longer trips than other companies

This question explores whether trip distance, a known predictor of tipping, acts as a confounding variable in the relationship between company and tip rate.

------------------------------------------------------------------------

### Initial exploration: Visual comparison

Let's look at the data. A boxplot allows us to compare the distribution of trip distances between Chicago Independents and all other companies:

```{r}
#| label: fu-distance-comparison
#| fig-cap: "Trip distance distribution by company group"

taxi |> 
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |> 
  filter(!is.na(distance), distance > 0) |> 
  ggplot(aes(x = company_group, y = distance, fill = company_group)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Do Chicago Infdependents perform longer trips?",
    subtitle = "Trip distance comparison",
    x = "",
    y = "Distance (miles)",
    fill = "Company group"
  ) +
  theme_minimal()
```

The boxplots show **substantial overlap** between the two groups. The median distances appear similar, and the interquartile ranges largely coincide. This visual evidence suggests that any difference in trip distances, if present, is likely small.

------------------------------------------------------------------------

#### Summary Statistics

```{r}
#| label: fu-distance-summary

distance_summary <- taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  group_by(company_group) |>
  summarise(
    mean_distance = mean(distance, na.rm = TRUE),
    median_distance = median(distance, na.rm = TRUE),
    sd_distance = sd(distance, na.rm = TRUE),
    q25 = quantile(distance, 0.25, na.rm = TRUE),
    q75 = quantile(distance, 0.75, na.rm = TRUE),
    n = n()
  )

distance_summary |>
  knitr::kable(digits = 2, caption = "Trip distance statistics by company group")
```

The Chicago Independents show a slightly higher mean distance, but the **medians are close**. The similarity in medians and quartiles suggests that the distribution shapes are comparable.

------------------------------------------------------------------------

#### Statistical comparison

While a t-test is often used to compare means, the bimodal distribution of trip distances (observed in Question 1) violates the normality assumption. A more robust approach is to compare distributions visually and examine effect size rather than rely solely on p-values.

```{r}
#| label: fu-distance-density
#| fig-cap: "Distance density by company group"

taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  filter(!is.na(distance), distance > 0, distance <25) |>
  ggplot(aes(x = distance, color = company_group, fill = company_group)) +
  geom_density(alpha = 0.3) +
  labs(
    title = "Trip distance density",
    subtitle = "Both groups show similar bimodal patterns",
    x = "Distance (miles)",
    y = "Density",
    color = "Company group",
    fill = "Company group"
  ) +
  theme_minimal()
```

The density plots reveal that both groups exhibit the same bimodal pattern observed in Question 1: a peak around 2 miles and another around 18 miles. The shapes are nearly identical, with only minor shifts in the relative heights of the peaks.

------------------------------------------------------------------------

### Statistical test

```{r}
# Run t-test and store results
t_test_result <- taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  with(t.test(distance ~ company_group))

# Extract components into a tidy dataframe
t_test_table <- data.frame(
  Statistic = c("t-value", "Degrees of freedom", "p-value", 
                "Mean difference", "CI lower", "CI upper"),
  Value = c(
    round(t_test_result$statistic, 3),
    round(t_test_result$parameter, 1),
    format.pval(t_test_result$p.value, digits = 3),
    round(diff(t_test_result$estimate), 2),
    round(t_test_result$conf.int[1], 2),
    round(t_test_result$conf.int[2], 2)
  )
)

# Display with kable
t_test_table |>
  knitr::kable(caption = "T-test results: Trip distance comparison")
```

------------------------------------------------------------------------

### Findings

The visual evidence leads to the following conclusion:

-   Boxplots show substantial overlap: the interquartile ranges and medians are very similar between groups

-   Density plots reveal identical bimodal structure: both groups have the same two peaks at approximately 2 and 18 miles

Any difference in means is small (less than 1 mile) and driven by slight shifts in the proportion of short vs. long trips, not by a systematic difference in trip lengths

Despite a statistically significant t-test (p = 0.001), the practical difference is negligible. With sample sizes in the thousands, even tiny differences can become "significant" while remaining meaningless in real-world terms.

------------------------------------------------------------------------

### Conclusion

**There is no meaningful difference in trip distances between Chicago Independents and other companies.**

The distributions are visually indistinguishable, with similar medians, quartiles, and overall shape. This finding has an important implication for our earlier analysis:

Since trip distances are essentially the same across company groups, the higher tip rate for Chicago Independents cannot be explained by longer trips.

The tip difference observed in Question 2 must therefore be attributed to other factors:

-   Service quality differences

-   Passenger demographics

-   Neighborhood characteristics

-   Payment method preferences

-   Other company-specific variables

This follow-up demonstrates the value of visual inspection over automated statistical tests, especially when working with non-normal distributions. The boxplot and density plotprovide a clearer answer than any p-value.
