---
title: "Project"
author: "Marco Sorbona"
subtitle: "Investigation into Chicago taxis"
echo: false
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

## Packages

```{r}
#| warning: true
#| message: true
library(tidyverse)
library(tidymodels)
library(scales)
```

## Overview of the data set

The `taxi` data set containing information on a subset of taxi trips in the city of Chicago in 2022.

Below is a snapshot of the data set showing the variables

```{r}
glimpse(taxi)
```

## Question 1

First question to answer is the following: *Is there there a relationship between the distance (in miles) someone travels in a taxi and if they tip or not?*

Rational: the question is asking if there is a relationship between the distance of the ride and the eventuality of getting a tip based on the distance.

-   Hp1: longer rides get a tip.

-   Hp2: shorter rides are cheaper so they get a tip

in this case we have to find a relationship between a `continuous variable` (miles) and a `categorical variable` (tip: Yes/No).

We can evaluate the presence of a relationship using graphical visualizations. the most appropriate in this case would be a bar plot having the tip on the "x" axes and the miles on the "y" axes.

## Question 2

## Your own question
