---
title: "Project"
author: "Marco Sorbona"
subtitle: "Investigation into Chicago taxis"
echo: false
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r}
#| warning: false
#| message: true
library(tidyverse)
library(tidymodels)
library(scales)
library(ggthemes)
```

## Data Overview

The `taxi` dataset contains information on a subset of taxi trips in Chicago during 2022.

Below is a snapshot of the dataset:

```{r}
#| label: glimpse
glimpse(taxi)
```

## Question 1: Is there a relationship between trip distance and tipping?

**Research question:** Do passengers tip differently based on how far they travel?

Hypotheses:

-   **Hp1**: Longer trips are more likely to receive a tip

-   **Hp2**: Shorter trips are cheaper, so passengers might tip more frequently

This question involves exploring the relationship between a **continuous variable** (distance in miles) and a **categorical variable** (tip: yes/no).

------------------------------------------------------------------------

### Initial Exploration: Tip Frequency

First, let's look at the overall proportion of trips that received tips:

#### Proportion with Bar plot

```{r}
#| label: q1-proportion-bar-plot
#| fig-cap: "Proportion of trips with and without tips"
taxi |>
  count(tip) |>
  mutate(perc = n / sum(n)) |>
  ggplot(aes(x = tip, y = perc, fill = tip)) +
  geom_col() +
  scale_y_continuous(labels = percent) +
  labs(
    title = "Tip proportion in Chicago taxi trips",
    subtitle = "The vast majority of trips receive tips",
    x = "Tip received",
    y = "Proportion of trips",
    fill = "Tip"
  ) +
  theme_minimal()
```

**Observation:** Most trips in this dataset received a tip.

------------------------------------------------------------------------

#### Summary Statistics

```{r}
#| label: summary-stat
taxi |>
  group_by(tip) |>
  summarise(
    mean_dist = mean(distance, na.rm = TRUE),
    median_dist = median(distance, na.rm = TRUE),
    sd_dist = sd(distance, na.rm = TRUE),
    n = n()
  )
```

------------------------------------------------------------------------

#### Distribution of Trip Distances by Tip Status

#### Histogram

```{r}
#| label: q1-histogram
#| fig-cap: "Distribution of trips with and without tips"
#| warning: false
ggplot(taxi, aes(x = distance, fill = tip)) +
  geom_histogram(binwidth = 0.15, alpha = 0.7, position = "identity") +
  scale_x_continuous(limits = c(-0.5, 25), breaks = seq(0,50,5)) +
  scale_y_continuous(breaks = seq(0, 800, 200)) +
  labs(
    title = "Distribution of trip distance by tip status",
    subtitle = "Multimodal distribution with peaks at ~2, ~12, and ~18 miles",
    x = "Distance (miles)",
    y = "Number of trips",
    fill = "Tip"
  ) +
  theme_minimal()
```

The histogram reveals a **multimodal distribution** with **four distinct peaks**:

-   **Peak 1: \~0 miles**. Very short/zero-distance trips

-   **Peak 2: \~2 miles**. Most common trip length

-   **Peak 3: \~12 miles**. Medium-length trips

-   **Peak 4: \~18 miles**. Long trips

#### What Could the 0-Mile Peak Represent?

The peak at 0 miles suggests data quality issues (perhaps trips with missing distance), which should be investigated or filtered.

**Possible explanations:**

1.  **Data errors**: Trips recorded with zero distance (technical issues?)

2.  **Cancelled trips**: Passenger cancelled but driver still recorded?

3.  **Very short hops**: Around the block, between nearby buildings

4.  **Airport waiting**: Taxis waiting in queue with zero movement

------------------------------------------------------------------------

#### Boxplot

```{r}
#| label: q1-boxplot
#| fig-cap: "Trip distance distribution by tipping status"
ggplot(taxi, aes(x = tip, y = distance, fill = tip)) +
  geom_boxplot() +
  labs(
    title = "Trip distance distribution by tip status",
    subtitle = "Boxplot shows many outliers at longer distances",
    x = "Tip received",
    y = "Distance (miles)",
    fill = "Tip"
  ) +
  theme_minimal()
```

The boxplot confirms the presence of numerous outliers at longer distances, making it less ideal for visualizing the overall pattern.

------------------------------------------------------------------------

#### Density Plot (Best Visualisation)

```{r}
#| label: q1-density-plot
#| fig-cap: "Density of trip distances by tipping status"
#| warning: false
ggplot(taxi, aes(x = distance, color = tip, fill = tip)) +
  geom_density(alpha = 0.4, linewidth = 0.5) +
  scale_x_continuous(limits = c(0, 25)) +
  labs(
    title = "Trip distance density by tip status",
    subtitle = "Longer trips show higher probability of tipping",
    x = "Distance (miles)",
    y = "Density",
    color = "Tip",
    fill = "Tip"
  ) +
  theme_minimal()
```

### Findings

The density plot reveals several patterns:

1.  **Short trips (\<5 miles)**: Approximately equal density for tipped and non-tipped trips. Passengers are equally likely to tip or not.

2.  **Medium trips (10-15 miles)**: A small peak shows non-tipped trips slightly more common.

3.  **Long trips (\>15 miles)**: Tipped trips show higher density, suggesting passengers are more likely to tip on longer journeys.

4.  **Peak at 0 miles**: Likely data quality issue (trips with zero recorded distance) that should be investigated.

### Conclusion

**There is a clear relationship between trip distance and tipping behavior.**

The density plot reveals that:

1.  **Short trips** (around 2 miles) have roughly equal probability of receiving a tip or not

2.  **Medium trips** (around 12 miles) show a higher probability of **not** receiving a tip

3.  **Long trips** (around 18 miles) show a substantially higher probability of receiving a tip

This pattern supports **Hypothesis 1**: longer trips are more likely to receive a tip. The relationship is not simply linear, there appears to be a threshold around 15 miles where tipping probability increases noticeably.

The density plot is the most effective visualization for this question as it clearly shows probability differences across the distance spectrum

## Question 2: Do taxi passengers tend to tip more for the company Chicago Independents than the other companies?

**Research question**: Do passengers tip at higher rate when riding with Chicago Independents compared to other taxi companies?

**Hypotheses**:

-   Hp0: There is no difference in tip rates between Chicago Independents and other companies

-   Hp1: Chicago Independents receive tips at a higher rate than other companies

This question involves exploring the relationship between a categorical variable (company) and the proportion of trips that received tips.

------------------------------------------------------------------------

### Initial exploration: Calculate tip rate

First, we need to calculate the tip rate for each company: the proportion of trips where a tip was given.

```{r}
#| label: q2-company-tip-rate
#| fig-cap: "Tip rate by company"

company_tip_rates <- taxi |>
  group_by(company) |>
  summarise(
    total_trips = n(),
    tipped_trips = sum(tip == "yes", na.rm = TRUE),
    tip_rate = tipped_trips / total_trips
  ) |>
arrange(desc(tip_rate))

company_tip_rates
```

------------------------------------------------------------------------

#### Highlight Chicago Independents

To make comparison easier, we create a flag for Chicago Independents

```{r}
company_tip_rates <- company_tip_rates |>
  mutate(
    highlight = if_else(company == "Chicago Independents", "CI", "Other")
  )

company_tip_rates
```

------------------------------------------------------------------------

#### Visualisation

A bar plot allows us to compare tip rates across companies, with Chicago Independents highlighted:

```{r}
ggplot(company_tip_rates, aes(x = reorder(company, tip_rate), y = tip_rate, fill = highlight)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = c("CI" = "red", "Other" = "gray")) +
  labs(
    title = "Do passengers tip more for Chicago Independents?",
    subtitle = "Tip rate = % of trips with a tip",
    x = "Company",
    y = "Tip rate",
    caption = "Chicago Independents highlighted in red"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

### Findings

The analysis shows that **Chicago Independents have the highest tip rate** among all companies, at approximately **95%** of trips receiving a tip. This is higher than the tip rate across other companies.

------------------------------------------------------------------------

### Discussion

The higher tip rate for Chicago Independents could be explained by several factors:

-   Better service quality: Drivers may provide a superior experience

-   Operational differences: Perhaps they serve different neighborhoods or longer trips

-   Passenger demographics: Their customer base might have different tipping norms

##### Limitations and follow-up questions

The current analysis cannot explain why Chicago Independents have higher tip rates. Access to more granular data would help investigate:

-   Do they operate primarily in wealthier neighborhoods?

-   Do they perform longer trips on average? (Previous analysis showed longer trips are tipped more frequently)

-   Is there variation in service quality (vehicle condition, driver professionalism)?

-   Do they use different payment methods that encourage tipping?

These questions point to the need for multivariate analysis to disentangle the factors contributing to higher tip rates.

------------------------------------------------------------------------

## Follow-up Question: Do Chicago Independents perform longer trips on average?

#### Rationale

Question 1 established that longer trips are more likely to receive tips. Question 2 found that Chicago Independents have the highest tip rate among all companies (approximately 95%). This follow-up investigates whether the higher tip rate for Chicago Independents might be explained by systematically longer trips, rather than company-specific factors like service quality.

#### Hypotheses

-   H0: There is no meaningful difference in trip distances between Chicago Independents and other companies

-   H1: Chicago Independents perform significantly longer trips than other companies

This question explores whether trip distance, a known predictor of tipping, acts as a confounding variable in the relationship between company and tip rate.

------------------------------------------------------------------------

### Initial exploration: Visual comparison

Let's look at the data. A boxplot allows us to compare the distribution of trip distances between Chicago Independents and all other companies:

```{r}
#| label: fu-distance-comparison
#| fig-cap: "Trip distance distribution by company group"

taxi |> 
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |> 
  filter(!is.na(distance), distance > 0) |> 
  ggplot(aes(x = company_group, y = distance, fill = company_group)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Do Chicago Infdependents perform longer trips?",
    subtitle = "Trip distance comparison",
    x = "",
    y = "Distance (miles)",
    fill = "Company group"
  ) +
  theme_minimal()
```

The boxplots show **substantial overlap** between the two groups. The median distances appear similar, and the interquartile ranges largely coincide. This visual evidence suggests that any difference in trip distances, if present, is likely small.

------------------------------------------------------------------------

#### Summary Statistics

```{r}
#| label: fu-distance-summary

distance_summary <- taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  group_by(company_group) |>
  summarise(
    mean_distance = mean(distance, na.rm = TRUE),
    median_distance = median(distance, na.rm = TRUE),
    sd_distance = sd(distance, na.rm = TRUE),
    q25 = quantile(distance, 0.25, na.rm = TRUE),
    q75 = quantile(distance, 0.75, na.rm = TRUE),
    n = n()
  )

distance_summary |>
  knitr::kable(digits = 2, caption = "Trip distance statistics by company group")
```

The Chicago Independents show a slightly higher mean distance, but the **medians are close**. The similarity in medians and quartiles suggests that the distribution shapes are comparable.

------------------------------------------------------------------------

#### Statistical comparison

While a t-test is often used to compare means, the bimodal distribution of trip distances (observed in Question 1) violates the normality assumption. A more robust approach is to compare distributions visually and examine effect size rather than rely solely on p-values.

```{r}
#| label: fu-distance-density
#| fig-cap: "Distance density by company group"

taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  filter(!is.na(distance), distance > 0, distance <25) |>
  ggplot(aes(x = distance, color = company_group, fill = company_group)) +
  geom_density(alpha = 0.3) +
  labs(
    title = "Trip distance density",
    subtitle = "Both groups show similar bimodal patterns",
    x = "Distance (miles)",
    y = "Density",
    color = "Company group",
    fill = "Company group"
  ) +
  theme_minimal()
```

The density plots reveal that both groups exhibit the same bimodal pattern observed in Question 1: a peak around 2 miles and another around 18 miles. The shapes are nearly identical, with only minor shifts in the relative heights of the peaks.

------------------------------------------------------------------------

### Statistical test

```{r}
# Run t-test and store results
t_test_result <- taxi |>
  mutate(
    company_group = if_else(company == "Chicago Independents", "CI", "Other")
  ) |>
  with(t.test(distance ~ company_group))

# Extract components into a tidy dataframe
t_test_table <- data.frame(
  Statistic = c("t-value", "Degrees of freedom", "p-value", 
                "Mean difference", "CI lower", "CI upper"),
  Value = c(
    round(t_test_result$statistic, 3),
    round(t_test_result$parameter, 1),
    format.pval(t_test_result$p.value, digits = 3),
    round(diff(t_test_result$estimate), 2),
    round(t_test_result$conf.int[1], 2),
    round(t_test_result$conf.int[2], 2)
  )
)

# Display with kable
t_test_table |>
  knitr::kable(caption = "T-test results: Trip distance comparison")
```

------------------------------------------------------------------------

### Findings

The visual evidence leads to the following conclusion:

-   Boxplots show substantial overlap: the interquartile ranges and medians are very similar between groups

-   Density plots reveal identical bimodal structure: both groups have the same two peaks at approximately 2 and 18 miles

Any difference in means is small (less than 1 mile) and driven by slight shifts in the proportion of short vs. long trips, not by a systematic difference in trip lengths

Despite a statistically significant t-test (p = 0.001), the practical difference is negligible. With sample sizes in the thousands, even tiny differences can become "significant" while remaining meaningless in real-world terms.

------------------------------------------------------------------------

### Conclusion

**There is no meaningful difference in trip distances between Chicago Independents and other companies.**

The distributions are visually indistinguishable, with similar medians, quartiles, and overall shape. This finding has an important implication for our earlier analysis:

Since trip distances are essentially the same across company groups, the higher tip rate for Chicago Independents cannot be explained by longer trips.

The tip difference observed in Question 2 must therefore be attributed to other factors:

-   Service quality differences

-   Passenger demographics

-   Neighborhood characteristics

-   Payment method preferences

-   Other company-specific variables

This follow-up demonstrates the value of visual inspection over automated statistical tests, especially when working with non-normal distributions. The boxplot and density plotprovide a clearer answer than any p-value.

---

## Question 3: Do tipping patterns vary by time of day or day of week? 

**Research question**: Do passengers tip at different rates depending on when they ride: specifically, by hour of day and day of week?

**Rationale**

Tipping behavior may be influenced by contextual factors that vary across time:

-   *Early morning*: Airport runs, commuters heading to work

-   *Midday*: Short trips for errands, lunch

-   *Evening*: Dinner out, social events, nightlife

-   *Late night*: Bar crowds, revelers (potentially more generous or less?)

-   *Weekends vs. weekdays*: Leisure vs. business travel

Understanding these patterns could help drivers optimize their shifts and provide insight into passenger psychology.

**Hypotheses**

-   H1: Tip rates vary significantly by hour of day

-   H2: Tip rates differ between weekdays and weekends

-   H3: There is an interaction between time of day and day of week (e.g., Friday evening differs from Monday evening)

---

### Initial exploration: 

```{r}
ggplot(taxi, aes(x = hour, fill = tip)) +
  geom_bar()
```

#### Tip rate by hour

```{r}
#| label: q3-tip-by-hour
#| fig-cap: "Tip rate by hour of day"

hourly_tips <- taxi |>
  group_by(hour) |>
  summarise(
    tip_rate = mean(tip == "yes", na.rm = TRUE),
    n_trips = n()
  )

# View the data
hourly_tips |> head(10)

# Plot
ggplot(hourly_tips, aes(x = hour, y = tip_rate)) +
  geom_line(color = "steelblue", linewidth = 0.5) +
  geom_point(aes(size = n_trips), color = "darkred", alpha = 0.7) +
  scale_y_continuous(labels = percent, limits = c(0.9, 1)) +  
  scale_x_continuous(breaks = seq(0, 23, 2)) +
  labs(
    title = "Tip rate by hour of day",
    subtitle = "Point size indicates number of trips",
    x = "Hour (24-hour clock)",
    y = "Tip rate",
    size = "Number of trips"
  ) +
  theme_minimal()
```

#### Tip rate by day of the week

```{r}
#| label: q3-tip-by-dow
#| fig-cap: "Tip rate by day of week"

dow_tips <- taxi |>
  group_by(dow) |>
  summarise(
    tip_rate = mean(tip == "yes", na.rm = TRUE),
    n_trips = n()
  )

# Plot (ensuring correct order)
dow_tips |>
  mutate(dow = factor(dow, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) |>
  ggplot(aes(x = dow, y = tip_rate, fill = dow)) +
  geom_col(alpha = 0.8) +
  geom_text(aes(label = paste0(round(tip_rate * 100, 1), "%"),
                y = tip_rate + 0.01), size = 3) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  labs(
    title = "Tip rate by day of week",
    x = "",
    y = "Tip rate"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

#### combine view

```{r}
#| label: q3-heatmap
#| fig-cap: "Tip rate by hour and day of week"

taxi |>
  group_by(dow, hour) |>
  summarise(
    tip_rate = mean(tip == "yes", na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  filter(n > 10) |>  # Remove sparse cells
  mutate(dow = factor(dow, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))) |>
  ggplot(aes(x = hour, y = dow, fill = tip_rate)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "#2166AC",      # Dark blue (low)
    mid = "#F7F7F7",      # White (mid)
    high = "#B2182B",     # Dark red (high)
    midpoint = 0.92, 
    labels = percent
  ) +
  scale_x_continuous(breaks = seq(0, 23, 3)) +
  labs(
    title = "Tip rate by hour and day of week",
    subtitle = "Blue = lower tip rates, red = higher",
    x = "Hour of day",
    y = "",
    fill = "Tip rate"
  ) +
  theme_minimal()
```

```{r}
#| label: q3-summary-table
#| tbl-cap: "Tip rates by time of day and weekend status"

taxi |>
  mutate(
    time_category = case_when(
      hour < 6 ~ "Late night (0-5)",
      hour < 12 ~ "Morning (6-11)",
      hour < 18 ~ "Afternoon (12-17)",
      TRUE ~ "Evening (18-23)"
    ),
    weekend = if_else(dow %in% c("Sat", "Sun"), "Weekend", "Weekday")
  ) |>
  group_by(time_category, weekend) |>
  summarise(
    tip_rate = mean(tip == "yes", na.rm = TRUE),
    n_trips = n(),
    .groups = "drop"
  ) |>
  arrange(desc(tip_rate)) |>
  knitr::kable(digits = 3, caption = "Tip rates by time category and weekend/weekday")
```

plot the above:

```{r}
#| label: q3-summary-data

summary_data <- taxi |>
  mutate(
    time_category = case_when(
      hour < 6 ~ "Late night (0-5)",
      hour < 12 ~ "Morning (6-11)",
      hour < 18 ~ "Afternoon (12-17)",
      TRUE ~ "Evening (18-23)"
    ),
    weekend = if_else(dow %in% c("Sat", "Sun"), "Weekend", "Weekday")
  ) |>
  group_by(time_category, weekend) |>
  summarise(
    tip_rate = mean(tip == "yes", na.rm = TRUE),
    n_trips = n(),
    .groups = "drop"
  )

# Ensure time_category is in correct order
summary_data <- summary_data |>
  mutate(
    time_category = factor(time_category, 
                           levels = c("Late night (0-5)", "Morning (6-11)", 
                                      "Afternoon (12-17)", "Evening (18-23)"))
  )
```

plot1:

```{r}
#| label: q3-summary-bar
#| fig-cap: "Tip rates by time of day and weekend status"

ggplot(summary_data, aes(x = time_category, y = tip_rate, fill = weekend)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = paste0(round(tip_rate * 100, 1), "%")),
            position = position_dodge(width = 0.7),
            vjust = -0.5, size = 3) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("Weekday" = "#1f77b4", "Weekend" = "#ff7f0e")) +
  labs(
    title = "Tip rates by time of day",
    subtitle = "Weekend vs. weekday comparison",
    x = "",
    y = "Tip rate",
    fill = "",
    caption = "Bars show tip rate; labels show exact percentage"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

plot2:

```{r}
#| label: q3-summary-facet
#| fig-cap: "Tip rates by time period, faceted by weekend/weekday"

ggplot(summary_data, aes(x = time_category, y = tip_rate, fill = weekend)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(tip_rate * 100, 1), "%")),
            vjust = -0.5, size = 3) +
  facet_wrap(~weekend) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("Weekday" = "#1f77b4", "Weekend" = "#ff7f0e")) +
  labs(
    title = "Tip rates by time of day",
    subtitle = "Weekday (left) vs. Weekend (right)",
    x = "",
    y = "Tip rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

plot3:

```{r}
#| label: q3-summary-dots
#| fig-cap: "Tip rates with confidence based on sample size"

ggplot(summary_data, aes(x = time_category, y = tip_rate, color = weekend)) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = tip_rate - 1.96 * sqrt(tip_rate * (1 - tip_rate) / n_trips),
                    ymax = tip_rate + 1.96 * sqrt(tip_rate * (1 - tip_rate) / n_trips)),
                width = 0.2, position = position_dodge(width = 0.5)) +
  scale_y_continuous(labels = percent, limits = c(0.8, 1)) +
  scale_color_manual(values = c("Weekday" = "#1f77b4", "Weekend" = "#ff7f0e")) +
  labs(
    title = "Tip rates by time of day",
    subtitle = "Error bars show 95% confidence intervals (wider = less certain)",
    x = "",
    y = "Tip rate",
    color = ""
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

plot4:

```{r}
#| label: q3-summary-heatmap
#| fig-cap: "Tip rates heatmap: time of day vs. weekend status"

summary_data |>
  mutate(
    weekend = factor(weekend, levels = c("Weekday", "Weekend")),
    label = paste0(round(tip_rate * 100, 1), "%")
  ) |>
  ggplot(aes(x = weekend, y = time_category, fill = tip_rate)) +
  geom_tile() +
  geom_text(aes(label = label), color = "white", size = 4) +
  scale_fill_gradient(low = "#08519C", high = "#F7F7F7", labels = percent) +
  labs(
    title = "Tip rates by time of day and weekend status",
    x = "",
    y = "",
    fill = "Tip rate"
  ) +
  theme_minimal()
```

plot5:

```{r}
#| label: q3-summary-clean-with-errors
#| fig-cap: "Tip rates by time of day (weekend vs. weekday) with 95% confidence intervals"

# First, calculate standard errors for each group
summary_data <- summary_data |>
  mutate(
    se = sqrt(tip_rate * (1 - tip_rate) / n_trips),  # Standard error
    margin = 1.96 * se,                               # 95% CI margin
    ci_lower = tip_rate - margin,
    ci_upper = tip_rate + margin
  )

# Now plot with error bars
ggplot(summary_data, aes(x = time_category, y = tip_rate, fill = weekend)) +
  geom_col(position = position_dodge(width = 0.6), width = 0.6, alpha = 0.9) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    position = position_dodge(width = 0.6),
    width = 0.2,
    linewidth = 0.5
  ) +
  scale_y_continuous(labels = percent, limits = c(0, 1), expand = c(0, 0)) +
  scale_fill_manual(values = c("Weekday" = "#2c7fb8", "Weekend" = "#f03b20")) +
  labs(
    title = "Do weekends have higher tip rates?",
    subtitle = "Tip rates by time of day, comparing weekday vs. weekend (error bars = 95% CI)",
    x = "",
    y = "Tip rate",
    fill = "",
    caption = paste("Based on", nrow(taxi), "Chicago taxi trips (2022)")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "top",
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )
```

```{r}
#| label: q3-honest-plot
#| fig-cap: "Tip rates by time of day (no meaningful pattern)"

summary_data |>
  mutate(
    # Flag low sample sizes
    reliable = if_else(n_trips > 500, "Reliable", "Low sample"),
    time_category = factor(time_category, 
                           levels = c("Late night (0-5)", "Morning (6-11)", 
                                      "Afternoon (12-17)", "Evening (18-23)"))
  ) |>
  ggplot(aes(x = time_category, y = tip_rate, color = weekend, shape = reliable)) +
  geom_point(size = 4, position = position_dodge(width = 0.5)) +
  geom_hline(yintercept = 0.92, linetype = "dashed", alpha = 0.5) +
  scale_y_continuous(labels = percent, limits = c(0.85, 1)) +
  labs(
    title = "Tip rates are consistently around 92%",
    subtitle = "No meaningful variation by time or day (dashed line = overall average)",
    x = "",
    y = "Tip rate",
    color = "",
    shape = "Reliability",
    caption = "Points with low sample size (<500 trips) are less reliable"
  ) +
  theme_minimal()
```

test\]
